The Galactic Chemical Evolution of Carbon: Implications for Stellar Nucleosynthesis

### 2024-08-29 carbon_paper/tex/main.tex
updated text (created 3panel degeneracy figure)

@@

\newcommand{\citetjack}{\citet{jack}}
\newcommand{\citepjack}{(Roberts et al.~2024, in prep.)}
\newcommand{\citealtjack}{Robert et al.~2024, in prep.}
\newcommand{\cxi}{\texttt{\hyperlink{C11}{C11}}}
\newcommand{\pxvi}{\texttt{\hyperlink{P16}{P16}}}
\newcommand{\kxvi}{\texttt{\hyperlink{K16}{K16}}}
\newcommand{\vxiii}{\texttt{\hyperlink{V13}{V13}}}
\newcommand{\cfactor}{2.4}
\newcommand{\nsubgiants}{16,365}

@@

% other
\newcommand{\ycmg}{\ensuremath{2.7 + 32\left(Z-\Zo\right)}}
\newcommand{\ycccmg}{1}
\newcommand{\yco}{1}
\newcommand{\fmeas}{20\%}

@@

\newcommand{\Yct}{\ensuremath{y_{\rm C}}}
\newcommand{\Ycc}{\ensuremath{y_{\rm C}^{\rm CC}}}
\newcommand{\Yoc}{\ensuremath{y_{\rm Mg}^{\rm CC}}}
\newcommand{\Ycagb}{\ensuremath{y_{\rm C}^{\rm AGB}}}
% \newcommand{\y}{\ensuremath{\rotatebox[origin=B,y=0.5ex]{180}{y}}}
\newcommand{\y}{Y}

\newcommand{\Mo}{%
    \ifmmode {\rm M}_{\sun}%
    \else {M$_{\sun}$}%
    \fi}

\newcommand{\Zo}{%
    \ifmmode Z_{\sun}%
    \else $Z_{\sun}$%
    \fi}

@@

@@

\newcommand{\dbstrike}[1]{{\color{Thistle} \sout{#1}}}
\newcommand{\dbadd}[1]{{\color{Thistle} #1}}
\newcommand{\dbnote}[1]{{\color{Thistle} \textit{\small (DB: #1)}}}

@@

Despite being one of the most abundant metals in the universe~\citep[e.g.][]{asplund+09}, C nucleosynthesis is poorly understood.
Formed in the cores of stars during He fusion, C is the first \dbadd{primary? element} after He. Additionally, C is one of the only light elements formed in low-mass stars \citep[e.g.,][]{KL14}. \dbnote{I don't like how many e.g.'s there are, and is this the right citation here?}
\dbadd{
Because the effects of first dredge up are mass-dependent, C and N abundances are frequently used as age indicators for RGB stars \citep{MG15, martig16, hasselquist19, vincenzo+21}. 
}
\dbnote{C abundnces for stellar structure (opacities) and star formation (line cooling)}
Broadly, it is well understood that both lower-intermediate-mass and high-mass stars produce C. However, the relative yields from these two classes of stars and their dependence on  metallicitys unknown \add{(see discussion below)}.

% \footnotetext{\strike{By metallicity, we mean the (mass) fraction of any element which is not H or He, denoted by $Z$. For the sun, we take $Z_\odot=0.016$. CHECK THIS~} }

Like many other metals, C nucleosynthesis \dbstrike{is poorly understood because its predicted yields} \dbadd{predictions from stellar evolution are highly uncertain \citep[see reviews][]{romano+10, KL14}}.
\strike{Yield predictions} are shaped by poorly understood processes, such as mass loss rates \citep{ventura+13}?, nuclear reaction networks, rotational mixing \citep{frischknecht+16, LC18}, and convection ~\citep{chieffi2001}. \dbnote{check citations here...}. 
Binarity (and the metallicity dependence thereof), e.g. MKB2019, farme,
Small changes in the initial mass ($0.01\Mo$) of massive stars can dramatically change the subsequent evolution \citep{bruenn+2023, emily+21}. As such, accurate yield predictions would require a much more densly sampled set of models in mass, metallicity, and rotation rates then are presently available. Furthermore, binary evolution is rarely included in massive star models with possibly important implications for yields \citep{farmer+21}, despite the large (metallicity-dependent) fraction of massive stars which live in binary systems \citep{MKB2019}.

\dbnote{added paragraph here}
In light of these difficulties with accurate theoretical modeling, recent work has instead sought to characterize chemical yields empirically.
\citet{HLA20} measured the abundance ratios in the ejecta of Cassiopeia A using spatial resolved X-ray spectra.
\citet{weinberg+19, weinberg+22} analyzed the abundance ratio trends of high- and low-alpha sequence stars in APOGEE~\citep{apogee17} to infer relative yields of prompt and delayed enrichment sources and the metallicity dependence thereof.
\citet{emily+19, emily+22, emily+23} applied the same methodology to GALAH~\citep{DeSilva2015, Martell2017}.
\citet{rodriguez+21, rodriguez+23} measured the mean Fe yield of core-collapse supernovae (CCSNe) using the radioactive tails of their lightcurves.

\dbnote{combine this with previous, shorten J23 description and call out other ongoing work e.g. Miquela's He, James's chemical equilibrium, ...}
Most notably for our purposes,~\citet{james+23} used Galactic chemical evolution (GCE) models to show that the correlation between gas-phase N and O abundances (see their Fig. 1) can be explained by the metallicity dependence of relative N and O yields. 
Variations in assumptions regarding GCE parameters, such as the detailed star formation history (SFH), did not impact their conclusions and are instead a likely source of scatter in the observed relation.
Here, we apply similar methodology to constrain C yields, which are closely related to N.

GCE connects a galaxy's evolutionary history to the detailed elemental abundances in its gas and stars.
Each enrichment process has characteristic chemical signatures and timescales, enabling us to reconstruct \strike{chemical} \add{enrichment} histories.
\add{As an element that is well-measured in stellar spectra, many} \dbadd{with large samples of stellar C abundancess}, \strike{Many} previous GCE works \add{have} focused on C abundances (\citealt{DTS78, BF06, prantzos+18, berg+19}; see also the recent reviews by \citealt{romano22} and \citealt{RM21}).
To date there is no consensus on the detailed origin of C; some authors have argued that high-mass stars should dominate~\citep[e.g.][]{prantzos+94, HEK00, romano+20, franchini+20, gustafsson22}, and others have argued that AGB stars should contribute at least half of the C in the universe~\citep[e.g.][]{tinsley79, chiappini+03, mattsson10, KKU11, rybizki+17, KKL20}.

    As our primary observational constraint, we use a sample of subgiant stars from the \apogee{} \citep{apogee17} selected by the criteria in \citetjack.
According to stellar evolution theory and observations, subgiants most accurately represent their birth composition \citep{gilroy89, korn+07, lind+08, souto+18, souto19}.
Fig.~\ref{fig:subgiants} shows the subgiant sample plotted in \caah\ and \caafe{}.\footnotemark{} In this sample, [C/Mg] increases with metallicity, and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. 
From observations of nearby stars and (extragalactic) gas, we know that [C/O] against [O/H] traces a banana shape (see Fig.~\ref{fig:gas_phase} and section~\ref{sec:gas}).
At the very lowest metallicities\strike{, where} \add{(}$[{\rm O/H}]\lesssim-2$\add{)},%
%
observations \strike{suggest} \add{indicate} that [C/Mg] declines with metallicity.
From $-2 \lesssim {\rm [O/H]}\lesssim -1$, [C/Mg] is roughly constant with metallicity. 
At higher metallicities ([O/H]~$\gtrsim -1$), [C/Mg] increases with metallicity.

\strike{Using the subgiant abundance trends, we will develop a model of the enrichment sources and evolution of C.}
\add{We will show in section~\ref{sec:results} below that these correlations in the bulk population abundances alone are quite constraining for the origin of C.}

\footnotetext{In this paper, we use the standard notation for chemical abundances. $[A/B] = \log_{10}\left(A/B\right) - \log_{10}\left(A_{\sun}/B_{\sun}\right)$, i.e. $[A/B]$ is the logarithm of the mass ratio between A and B, scaled such that $[A/B]=0$ for the sun (see Table~\ref{tab:fiducial_mod} for solar scale.)}

@@

\section{Data Selection}\label{sec:data_selection}

\begin{figure*}
    \centering
    \includegraphics{subgiants.pdf}
    \caption{The [C/Mg] ratio against [Mg/H] (top) and [Mg/Fe] (bottom) for the \citetjack~sample of \apogee{} subgiants. On the top, we plot high and low-$\alpha$ stars in blue and orange, using the separation defined in Equation \ref{eq:high_alpha} (the high and low-$\alpha$ stars are named for their high or low $\alpha$-element to Fe ratios, or in this case, Mg/Fe). On the bottom, we colour-code stars according to their [Mg/H] abundance.} \label{fig:subgiants}
\end{figure*}

% Subgiants are a useful empirical benchmark because their photospheric abundances should reflect their birth mixture. 
As an emperical benchmark, we use subgiant stars, whose photospheres accurately represent birth abundances. 
During main sequence evolution, metals can fall to below the convective envelope (i.e., ``gravitational settling").
When stars leave the main sequence and become subgiants, these metals are mixed with the convective envelope \citep[]{gratton+00, souto19}. 
 \textit{First dredge up} during the red giant phase brings CNO processed core material to the surface \citet{iben67, KL14}.

We use the sample of \apogee\ DR17 subgiants from \citetjack, selecting stars within the $\log g$-$T_\text{eff}$ polygon given by
 \begin{equation}
    \begin{cases} \label{eq:subgiant_selection}
        \log \text{g} \geq 3.5 \\
        \log \text{g} \leq 0.004\,T_{\rm eff} - 15.7 \\
        \log \text{g} \leq 0.000706\,T_{\rm eff} + 0.36 \\
        \log \text{g} \leq -0.0015\,T_{\rm eff} + 12.05 \\
        \log \text{g} \geq 0.0012\,T_{\rm eff} - 2.8. \\
    \end{cases}
\end{equation}
Also following \citet{jack}, we exclude stars with the flags:
        \verb|APOGEE_MIRCLUSTER_STAR|,
        \verb|APOGEE_EMISSION_STAR|,
        \verb|APOGEE_EMBEDDEDCLUSTER_STAR|,
        \verb|APOGEE2_YOUNG_CLUSTER|,
            \verb|APOGEE2_W345|,
        \verb|APOGEE2_EB|, and
        \verb|ASPCAP_BAD|.
We additionally remove stars with no reported C, Mg, or Fe abundances. These cuts yield a sample containing \nsubgiants\ subgiants.

@@

Table \ref{tab:fiducial_mod} contains our fiducial yields and abundance scale. 
For O, Mg, and Fe, we adopt the yields in \citet{david_fe}. We use a scaled variation of the N yield from \citet{james+23}. 
Following \citet{james+21, james+23}, we take the \ia{} delay time distribution to be a $t^{-1.1}$ power-law with a minimum delay time of 150\,Myr, as suggested by the observations of \citet{maoz+12}. We use solar abundances from \citet{magg+22} with a +0.04 gravitational settiling correction as in \citet{david_fe}. \dbadd{Magg22 finds Z/X = 0.0225 for the surface? what do we use for the solar metallicity?}. Assuming He abundance from $Y=0.2725$ from \citet{AAG2021}, resulting in a $Z_\odot=0.01606$

\begin{table}
	\centering
    \caption[]{Solar abundance scale and fiducial yields (in units of SSP~birth mass). See section \ref{sec:agb} for the definition of \cxi. The solar abundance scale is \citet{magg+22} + 0.04. }
	\label{tab:fiducial_mod}

	\begin{tabular}{l l l l l}
		\hline
        Element & $Z_{X,\,\sun}$ & $y_X^{\rm cc}$ & $\y_X^{\rm agb}$ & $y_X^{\rm ia}$  \\
		\hline
        C & 0.00339 & Eq.~\ref{eq:y_cc} & $\cfactor\times$\cxi &  0 \\
        O & 0.00733 & 7.12e-3 & 0 & 0 \\
        Mg & 0.000671 & 6.69e-4 & 0 & 0 \\
        Fe & 0.00137 & 4.72e-4 & 0 & 6.69e-4 \\
        N &0.00104 & 4.001e-4 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
        % C & 0.00339 & Eq.~\ref{eq:zeta} & $\cfactor\times$\cxi &  0 \\
        % O & 0.00733 & 0.015 & 0 & 0 \\
        % Mg & 0.000671 & 0.00185 & 0 & 0 \\
        % Fe & 0.00137 & 0.0012 & 0 & 0.00214 \\
        % N &0.00104 & 0.00072 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

\dbnote{I like the definition of Y to go here since that is model independent, but IMF-integrated could still be specific to AGB/CC.}

@@

An Asymptotic Giant Branch (AGB) star is a low mass star during its last nuclear burning phase (He shell burning; see e.g. \citealt{PR2023}) which enriches the ISM through mass loss during thermal pulses.
    In this work, we explore four different stellar AGB yield tables from the literature which provide
    well-sampled grids in mass and metallicity. We refer to the tables as 
\begin{description}
    \item \hypertarget{C11}{\texttt{C11}}: \citet{cristallo+11, cristallo+15}
    \item \hypertarget{V13}{\texttt{V13}}: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \hypertarget{K16}{\texttt{K16}}: \citet{KL16, karakas+18}
    \item \hypertarget{P16}{\texttt{P16}}: \citet{pignatari+16, ritter+18, battino+19, battino+21}
\end{description}
Table~\ref{tab:agb} notes the masses and metallicities for each grid of yields.
We use the same tables as \citet{james+23}, except that we have swapped \citet{karakas10} for \citet{ritter+18}.

We define a yield to be the fraction of a star's initial mass $M_{\rm birth}$ which is converted into freshly synthesized C. If $\Delta Z_{\rm C}$ represents the change in the mass fraction of C abundance, then
\begin{equation}
    \y_{\rm C}^{\rm AGB} =  \Delta Z_{\rm C} \frac{M_{\rm ejected}}{M_{\rm birth}}.
\end{equation}
The yields may be negative if the material returned to the ISM has a lower $Z_{\rm C}$ than the material the star was formed from.

Fig.~\ref{fig:y_agb} compares the stellar \agb\ C yields for these four models.
Most models agree on general yield trends in mass and metallicity.
C production peak between masses of $\sim$ 2--4 \Mo\ and declines as stars become more or less massive. As metallicity increases, the net yield decreases and the mass of peak C production increases slightly.

>>>>

\usepackage{tikz}

@@

\newcommand{\fruity}{\texttt{\hyperlink{fruity}{Fruity}}}
\newcommand{\nugrid}{\texttt{\hyperlink{nugrid}{NuGrid}}}
\newcommand{\monash}{\texttt{\hyperlink{monash}{Monash}}}
\newcommand{\aton}{\texttt{\hyperlink{aton}{ATON}}}
\newcommand{\cfactor}{1.44}
\newcommand{\nsubgiants}{14,069}

@@

% other
\newcommand{\ycmg}{\ensuremath{2.7 + 32\left(Z-\Zo\right)}}
\newcommand{\fmeas}{20\%}

@@

\newcommand{\Yct}{{y_{\rm C}}}
\newcommand{\Ycc}{{y_{\rm C}^{\rm CC}}}
\newcommand{\Yoc}{{y_{\rm Mg}^{\rm CC}}}
\newcommand{\Ycagb}{{y_{\rm C}^{\rm AGB}}}
\newcommand{\aagb}{\alpha_{\rm C}^{\rm AGB}}
\newcommand{\zagb}{\zeta_{\rm C}^{\rm AGB}}
\newcommand{\zcc}{\zeta_{\rm C}^{\rm CC}}
 \newcommand{\yb}{\ensuremath{\rotatebox[origin=B,y=0.5ex]{180}{y}}}
\newcommand{\y}{Y}
\newcommand{\fagb}{f_{\rm C}^{\rm AGB}}

\newcommand{\Mo}{ {\rm M}_{\sun}}

\newcommand{\Zo}{ Z_{\sun}}

@@

% Caption macros
\newcommand{\captionline}[2][very
thick]{\tikz[baseline={([yshift=-.5ex]current bounding box.base)}]{
\draw[#2,#1, line cap=round] (0,0) -- (0.3,0); }}

@@

\newcommand{\dbstrike}[1]{{\color{Thistle} \sout{#1}}}
\newcommand{\dbadd}[1]{{\color{Thistle} #1}}
\newcommand{\dbnote}[1]{ {\color{Thistle} \textit{\small (DB: #1)}} }

@@

\begin{figure*}
    \centering
    \includegraphics{subgiants.pdf}
    \caption{The [C/Mg] ratio against [Mg/H] (left) and [Mg/Fe] (right) for the \citet{jack}~sample of \apogee{} subgiants. On the left, we plot high and low-$\alpha$ stars in blue and orange, using the separation defined in Equation \ref{eq:high_alpha} \dbstrike{(the high and low-$\alpha$ stars are named for their high or low $\alpha$-element (Mg) to Fe ratios)}. On the right, we colour-code stars according to their [Mg/H] abundance.} \label{fig:subgiants}
\end{figure*}

Despite being the most abundant metal in the Universe~\citep[e.g.][]{magg+22}, the nucleosynthetic origin of C is inadequately understood. \dbnote{need synonyms for poorly understood}
Broadly, studies agree that both high mass and asymptotic giant branch (AGB) stars produce C, but their relative yields are disputed.
From a Galactic chemical evolution (GCE) perspective, some authors have argued that high-mass stars should dominate~\citep[e.g.][]{prantzos+94, HEK00, romano+20, franchini+20, gustafsson22}, and others have argued that AGB stars should contribute at least half of the C in the universe~\citep[e.g.][]{tinsley79, chiappini+03, mattsson10, KKU11, rybizki+17, KKL20}. 
\dbadd{We aim to empirically characterize the yields of C and the ratio of AGB and CCSNe production.}

% Formed in the cores of stars during He fusion, C is the first \dbadd{primary? element} after He. Additionally, C is one of the only light elements formed in low-mass stars \citep[e.g.,][]{KL14}. \dbnote{I don't like how many e.g.'s there are, and is this the right citation here?}
%\dbadd{
% Because the effects of first dredge up are mass-dependent, C and N abundances are frequently used as age indicators for RGB stars \citep{MG15, martig16, hasselquist19, vincenzo+21}. 
% }
% \dbnote{C abundnces for stellar structure (opacities) and star formation (line cooling)}

Predictions of C yields from stellar evolution and supernova (SN) models are riddled with uncertainties (see, e.g., the reviews/summaries by \citealt{romano+10, KL14}).
Stellar yields are shaped by many poorly understood processes, such as mass loss rates \citep{sukhbold+16, beasor+2020}, rotational mixing \citep{frischknecht+16, LC18}, reaction rates \dbadd{citations}, and convection \citep{chieffi2001, ventura+13}.
\dbnote{I checked these, is sukhbold the best reference on mass loss?}
\dbadd{In addition, stellar parameters such as convection and mass loss are calibrated to match observed stars and cannot be predicted from first principles.} \dbnote{need to check if this means SE models are calibrated to match yields or just stellar parameters.}
Small changes in the initial mass of only $\sim$$0.01 M_\odot$  were recently shown to have dramatic effects on the evolution and explosion outcomes of massive stars \citep{bruenn+2023, vartanyan_burrows2023}.
Furthermore, many stars are found in binary systems \dbadd{(e.g. \citealt{MKB2019}, others?)}, the effects of which are almost entirely unexplored in stellar nucleosynthesis models \citet{farmer+21}.
Therefore, accurate yield predictions from stellar theory requires both a much greater understanding of stellar physics and a much more densely sampled grid of models than is currently available.

In light of the challenges of stellar theory, recent work has instead sought to characterize chemical yields empirically.
\citet{HLA20} measured the abundance ratios in the ejecta of Cassiopeia A using spatial resolved X-ray spectra.
\citet{weinberg+19, weinberg+22} analyzed the abundance ratio trends of high- and low-alpha sequence stars in APOGEE~\citep{apogee17} to infer relative yields of prompt and delayed enrichment sources and the metallicity dependence thereof.
\citet{emily+19, emily+22, emily+24} applied the same methodology to GALAH~\citep{DeSilva2015, Martell2017}.
\citet{rodriguez+21, rodriguez+23} measured the mean Fe yield of core-collapse supernovae (CCSNe) using the radioactive tails of their lightcurves.
\dbnote{even papers as far back as at least https://ui.adsabs.harvard.edu/abs/1997NuPhA.621...27G/abstract do this. Include this series as well: https://ui.adsabs.harvard.edu/abs/2023MNRAS.525.1329L/abstract}

% \note{Reading through at least at this moment, I'd say this is a reasonable place for a new paragraph. I wouldn't worry about referencing Weller et al. (2024) since He chemical evolution is not super relevant here.}
 Folowing many previous works  (e.g. \citealt{DTS78, BF06, prantzos+18, berg+19}; see also the recent reviews by \citealt{romano22} and \citealt{RM21}.) this paper focuses on the evolution of C in GCE models.
Most notably for our purposes, \citet{james+23} recently showed that the enrichment history of the Galactic disk can be described as an equilibrium phenomenon.
In equilibrium GCE, trends in abundance ratios with metallicity arise as a consequence of trends in yield ratios (see discussion in their section {\color{red} X.Y}).
\citet{james+23} had previously shown that the correlation between gas-phase N and O abundances (see their Fig. 1) could be described as a consequence of the metallicity dependence of the relative N and O yields.
The equilibrium scenario suggests that this argument should extend to any element whose enrichment by stellar populations does not require long timescales ($\gtrsim$few Gyr).
We therefore extend \citepos{james+23} arguments to C, which is closely related to N.

We use a sample of subgiant stars from APOGEE \citep{apogee17} as our primary observational constraint.
\dbadd{APOGEE Subgiants provide a large, homogenous sample of well-measured C abundances unaffected by stellar evolution}.
Subgiants advantageously have photospheric abundances directly representing birth C abundances \citep{gilroy89, korn+07, lind+08, souto+18, souto19}.
In other evolutionary phases, the measured abundances are known to be affected by internal processes (see discussion in section 2 below), namely gravitational settling in main sequence stars {\color{red} (reference)} and dredge-up in red giants {\color{red} (reference)}, 
Our sample should greatly mitigate these difficulties with inferring birth abundances of C, which have been a significant source of uncertainty for previous investigations of its nucleosynthetic origin.

% \dbnote{not sure the observations below need to be  too detailed here? Maybe just state that C/O appears to have a universal shape across environments and direct to that section since our main focus is on APOGEE? and disk chemistry?  }
% \dbstrike{
% From observations of nearby stars and (extragalactic) gas, we know that [C/O] against [O/H] traces a banana shape (see Fig.~\ref{fig:gas_phase} and section~\ref{sec:gas}).
% At the very lowest metallicities ($[{\rm O/H}]\lesssim-2$),%
% %
% observations indicate that [C/Mg] declines with metallicity.
% From $-2 \lesssim {\rm [O/H]}\lesssim -1$, [C/Mg] is roughly constant with metallicity. 
% At higher metallicities ([O/H]~$\gtrsim -1$), [C/Mg] increases with metallicity.
% }
% \dbstrike{We will show in section~\ref{sec:results} below that these correlations in the bulk population abundances alone are quite constraining for the origin of C.}

@@

\footnotetext{In this paper, we use the standard notation for chemical abundances. $[A/B] = \log_{10}\left(A/B\right) - \log_{10}\left(A_{\sun}/B_{\sun}\right)$, i.e. $[A/B]$ is the logarithm of the mass ratio between A and B, scaled such that $[A/B]=0$ for the sun (see Table~\ref{tab:fiducial_mod} for solar scale.) }

% \footnotetext{\strike{By metallicity, we mean the (mass) fraction of any element which is not H or He, denoted by $Z$. For the sun, we take $Z_\odot=0.016$.} }

\section{The Subgiant Sample}\label{sec:data_selection}

% Subgiants are a useful empirical benchmark because their photospheric abundances should reflect their birth mixture. 
As an emperical benchmark, we use subgiant stars, whose photospheres accurately represent birth abundances. 
During main sequence evolution, metals can fall to below the convective envelope (i.e., ``gravitational settling").
When stars leave the main sequence and become subgiants, these metals are mixed with the convective envelope \citep[]{gratton+00, souto19}. 
 \textit{First dredge up} during the red giant phase brings CNO processed core material to the surface \citep{iben67, KL14}.

We use the sample of \apogee\ DR17 subgiants from \citet{jack}, selecting stars from allStar-dr17-synspec\_rev1 within the $\log g$-$T_\text{eff}$ polygon given by
 \begin{equation}
    \begin{cases} \label{eq:subgiant_selection}
        \log \text{g} \geq 3.5 \\
        \log \text{g} \leq 0.004\,T_{\rm eff} - 15.7 \\
        \log \text{g} \leq 0.0007\,T_{\rm eff} + 0.36 \\
        \log \text{g} \leq -0.0015\,T_{\rm eff} + 12.05 \\
        \log \text{g} \geq 0.0012\,T_{\rm eff} - 2.8. \\
    \end{cases}
\end{equation}
Also following \citet{jack}, we exclude stars with the flags:
        \verb|APOGEE_MIRCLUSTER_STAR|,
        \verb|APOGEE_EMISSION_STAR|,
        \verb|APOGEE_EMBEDDEDCLUSTER_STAR|,
        \verb|APOGEE2_YOUNG_CLUSTER|,
        \verb|APOGEE2_W345|,
        \verb|APOGEE2_EB|, and
        \verb|DUPLICATE|.
We additionally remove stars with no reported C, Mg, or Fe abundances (in the {\tt X\_FE} columns). These cuts yield a sample containing \nsubgiants\ subgiants.

Fig.~\ref{fig:subgiants} shows the subgiant sample plotted in \caah\ and \caafe{}.\footnotemark{} In this sample, [C/Mg] increases with metallicity, and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H].

\dbadd{
Not shown here, we investigate the median trends for different samples and surveys, including APOGEE Giants with mixing corrections \citep{vincenzo+21}, APOGEE dwarfs, Gaia-ESO dwarfs, GALAH subgiants, LAMOST MRS (CITE), and smaller spectroscopic samples of stars. The predominant difference between any survey with our sample is a systematic offset in [C/Mg] in both [Mg/Fe] and [Fe/H] of at most 0.2 dex. The exception is GALAH, who's C abundances are highly incomplete, based on a single line and thus C measurements are likely unreliable (see e.g. Emily). If the median trend of [C/Mg] is offset from what we assume, then we can simply scale our C yields by the appropriate factor (e.g. -0.15 dex => $10^{-0.15} \approx 0.7$, so $y_{\rm C} \rightarrow  0.7 y_{\rm C}$) without significantly changing our results.
}

@@

Table \ref{tab:fiducial_mod} contains our fiducial yields and abundance scale. We use solar abundances from \citet{magg+22} with a +0.04 gravitational settling correction  and assuming $X_\odot= 0.71, \mu_{H} = 1$ as in \citet{david_fe}. We thus assume $Z_\odot=0.016$ as apposed to $Z=0.014$ in \citet{asplund+09}.
% Z = (1-Y) / (1 + X/Z)

\begin{table}
	\centering
    \caption[]{Solar abundance scale and fiducial yields (in units of SSP~birth mass). See section \ref{sec:agb} for the definition of \fruity. The solar abundance scale is \citet{magg+22} + 0.04. }
	\label{tab:fiducial_mod}

	\begin{tabular}{l l l l l}
		\hline
        Element & $Z_{X,\,\sun}$ & $y_X^{\rm cc}$ & $\y_X^{\rm agb}$ & $y_X^{\rm ia}$  \\
		\hline
        C & 0.00339 & Eq.~\ref{eq:y_cc} & $\cfactor\times$\fruity &  0 \\
        O & 0.00733 & 7.12e-3 & 0 & 0 \\
        Mg & 0.000671 & 6.52-4 & 0 & 0 \\
        Fe & 0.00137 & 4.72e-4 & 0 & 6.69e-4 \\
        N &0.00104 & 4.001e-4 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
        % C & 0.00339 & Eq.~\ref{eq:zeta} & $\cfactor\times$\fruity &  0 \\
        % O & 0.00733 & 0.015 & 0 & 0 \\
        % Mg & 0.000671 & 0.00185 & 0 & 0 \\
        % Fe & 0.00137 & 0.0012 & 0 & 0.00214 \\
        % N &0.00104 & 0.00072 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

We define a stellar yield as the fraction of a star's initial mass $M$ which is converted into freshly synthesized C. If $\Delta Z_{\rm C}$ represents the change in the mass fraction of C abundance, then the stellar yield is
\begin{equation}
    \y_{\rm C}^{\rm AGB} =  \Delta Z_{\rm C} \frac{M_{\rm ejected}}{M} = 
    \frac{W_{\rm C}(M) - Z_{\rm C}\, M_{\rm ejected}}{M}
\end{equation}
where $\Delta Z_{\rm C}$ represents the change in mean metallicity of the initial composition the star was born from, $M_{\rm ejected} = M - M_{\rm rem}$ is the total ejected wind mass through the stars lifetime, $W_{\rm C}$ is the total mass of C ejected through winds, and $Z_{\rm C}$ is the initial C mass fraction of the star. 
For CCSNe, 
\begin{equation} \label{eq:imf-cc-yield}
   Y_{\rm C}^{\rm CC}(Z) = 
    \frac{E(M)\,M_{\rm C}(M) + W_{\rm C}(M) - Z_{\rm C}\,M}{M}
\end{equation}
\dbnote{Is the - Z M term right or should M be Mzams - Mrem ?}
where $E(M)$ is the explodability function, $M_{\rm C}(M)$ is the mass ejected in a successful supernovae, and $W_{\rm C}$ and $Z_{\rm C}$ are as above.
We assume 8\,$\Mo$ is the minimum mass of \cc\ progeneters (see \VICE\ documentation and \citealt{emily+21}). 
Note that a yield may be negative if a star expels gas with lower $Z_{\rm C}$ than the star was born with.

An IMF-averaged yield adds together the yields of stars of each mass, weighted by the fraction of stars of each mass (the IMF). 
The IMF-averaged yield $y_X^{\rm proc}$ for element $X$ under the enrichment process ${\rm proc}$ is given by
\begin{equation} \label{eq:imf-agb-yield}
    y_{X}^{\rm proc}(Z,\tau) = 
    \frac{
    \int_{M_{\rm to}(\tau)}^{M_{\rm max}} 
    \y_{\rm X}^{\rm proc}(M, Z) M\ 
    \frac{dN}{dM}\ dM
}
{
    \int_{M_{\rm min}}^{M_{\rm max}}\ M\ \frac{dN}{dM}\ dM
}
\end{equation}
where ${dN}/{dM}$ is the IMF and $M_{\rm to}(\tau)$ is the mass of stars with lifetime $\tau$, and $M_{\rm min}=0.08$ and $M_{\rm max} = 100$ are the assumed maximum and minimum values of the IMF.
$y$ thus represents the amount of freshly synthesized material returned to the ISM per unit of star formation. 
Note that in detail, the core collapse enrichment occurs on such short timescales that $M_{\rm to} = M_{\rm min, cc} = 8\,\Mo$

@@

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}
        \caption[]{\dbnote{replace y label with stellar C yield}. The net fractional \agb\ C yield  plotted as a function of initial stellar mass $M$ and colour-coded according to metallicity. The black dashed line shows $\y=0$ for reference. Each panel represents yields from one of four \agb\ models: \fruity{}, \aton{}, \monash{}, \nugrid{} (see sections \ref{sec:agb}) }

    \caption[]{\textbf{Left} \dbnote{ylabel: integrated carbon AGB yield} The (IMF-weighted) \agb\ C yield $\Ycagb$ as a function of metallicity for each of the \agb\ yield models. ($\Ycagb$ is the net mass of C produced by \agb\ stars per unit mass of star formation, after 10\,Gyr and assuming a \citealt{kroupa01} IMF.)
    \textbf{Right} cumulative C return as a function of age for a solar metallicity single stellar population. The dashed black line shows the delay time distribution of type Ia supernovae ($\propto t^{-1.1}$) for comparison. The minimum of \aton{} is $y(t)/y(t_{\rm end})=-46$. \dbnote{I think this changed with the new integration scheme...}
}

An Asymptotic Giant Branch (AGB) star is a low mass star during its last nuclear (He shell) burning phase which enriches the interstellar medium (ISM) through mass loss during thermal pulses.(see e.g. \citealt{PR2023} \dbnote{or more introductory textbook like Ryden intro to astro?}) . 
In this work, we explore four different stellar AGB yield tables from the literature which provide well-sampled grids in mass and metallicity. We refer to the tables as 
\begin{description}
    \item \hypertarget{fruity}{\texttt{FRUITY}}: \citet{cristallo+11, cristallo+15}
    \item \hypertarget{aton}{\texttt{ATON}}: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \hypertarget{monash}{\texttt{Monash}}: \citet{KL16, karakas+18}
    \item \hypertarget{nugrid}{\texttt{NuGrid}}: \citet{pignatari+16, ritter+18, battino+19, battino+21}
\end{description}
Table~\ref{tab:agb} notes the masses and metallicities for each grid of yields.
We use the same tables as \citet{james+23}, except that we have swapped \citet{karakas10} for \citet{pignatari+16}.

Fig.~\ref{fig:y_agb} compares the stellar \agb\ C yields for these four models.
Most models agree on general yield trends in mass and metallicity.
C production peak between masses of $\sim$ 2--4 $\Mo$ and declines as stars become more or less massive. As metallicity increases, the net yield decreases and the mass of peak C production increases slightly.

In order to integrate yields across the metallicity and mass range of interest, we interpolate yield tables linearly in both $Z$ and mass. Below the last sampled mass, $Y_{\rm C}^{\rm AGB}$ is linearly interpolated to 0 at $0.5\Mo$. The choice of yields for stars of masses $\Mo \lesssim 1.5$ produces about a 20\% uncertainty in the late DTD (past 3\,Gyr) of \agb\ C production from these models. More complete grids in mass and metallicity would be needed to more accurately compare GCE predictions between AGB evolution models. Furthermore, note the non-monotonic IMF-averaged yields in \nugrid\ and \aton\ particularly at low metallicities.

The left panel of Fig.~\ref{fig:agb-ssp} shows IMF-averaged C yields for each \agb\ model as a function of metallicity.
The \agb\ models mostly differ in their yield normalization and metallicity dependence.  
The normalizations span a factor of \about{2}.
For example, at $Z = 0.1\Zo$, each models predicts $Y_{\rm C}^{\rm AGB}$ to be between 0.0008 and
0.0016.
To characterize the metallicity dependence of each table, we fit the yield to a linear model in $\log Z$
\begin{equation}\label{eq:agb_z_approx}
    y_{\rm C}^{\rm AGB}(Z) = y_0 + \zagb \log(Z / \Zo).
\end{equation}
Table \ref{tab:agb} contains the fit parameters $y_0$ and $\zagb$ for each AGB table\footnote{we perform a MCMC fit using standard RMS error to estimate best parameters and uncertanties.}
The metallicity dependence of the yields span a factor of \about{3}.
Variations in \agb\ model yields are due to different choices of reaction rates, convection treatments, and mass-loss rates \citet{ventura+15} \dbnote{better ref here, is there a good review on AGB physics uncertanties?}.
